\chapter*{Acknowledgements}

I would not be able to accomplish this work without the priceless contribute from a lot of people to whom I am grateful.

First of all, I would like to thank my family, specially my parents and my grandfather, for all the constant support and the help to find motivation when I needed the most, and, most importantly, for the opportunity that was given to me in order to choose this field of study in this faculty. Without them and their sacrifice, it would not be possible, and I'll never forget that. I sure hope to have seized that opportunity in the right way.

To my friends and colleagues, the most amazing group of people I have met, I cannot thank you enough for being there for me, even in the worst moments. It was an immense privilege to be at your side during this course. Thanks for giving me strength to always carry on during the course, and for all the countless help. Without you, I probably wouldn't be writing this now. And Eduardo, you'll never be forgotten!

To my supervisor, Prof. Teresa Galvão, for all the support and precious advice, for all the valuable suggestions and the availability in the difficult times.

To Eng. António Nunes, my co-supervisor, for always being there to help and contribute with his valuable input, giving great advice and ideas for this work's development. His curiosity inspired me to explore new types of possible solutions to overcome the challenged presented in this work. I surely hope to have meet his expectations, because this project is born from his idea and because I recognise the great potential of this project.

To all the people at CAS, for making me comfortable and providing a great place to work.

To OPT, thanks for all the expertise and valuable feedback given in the final testing phase, and all the demonstrated interest in the project.

To the users of Stack Overflow, who 'saved my life' several times during the implementation phase of this work. Without that platform as a valuable help tool, the amount of work accomplished would surely be less.

Finally, I would like to thank all the people that took some of their own time to participate in the focus group and the usability testing. Without them, this work would not be possible for sure.



\vspace{10mm}
\flushleft{Marco Amador}
